Consistency and variability in the development of children's drawings of object categories

### 2021-06-25 writing/Manuscript_June2021.Rmd
latest edits

### Automated recognition scores
We used a combination of deep CNN activations and logistic regressions to obtain automated recognition scores, as per our pre-registered protocol. To encode the high-level visual features of each sketch, we used the VGG-19 architecture \cite{simonyan2014very}, a deep convolutional neural network pre-trained on Imagenet classification. We used model activations in the second-to-last layer of this network, which is the first fully connected layer of the network (FC6). Raw feature representations in this layer consist of flat 4096-dimensional vectors, to which we applied channel-wise normalization across all filtered drawings in the dataset.  Next, we used these features to train object category decoders. We then trained a 12-way logistic classifier with L2 regularization (tolerance = .1, regularization = .1), and used this classifier to estimate the category label for each drawing in the dataset.To avoid any bias due to imbalance in the distribution of drawings over categories, we randomly under sampled such that there were an equal number of drawings for each combination of geographical site (San Jose, Beijing) and the 12 categories. No additional metadata about the age of the child who produced each sketch was provided to the decoder. This procedure was repeated for each drawing in the dataset, yielding a binary a recognition score for each drawing.

@@

This anlaysis plan was pre-registered at https://osf.io/qymjr/; however, we had initially planned to use automated recognition scores as our main dependent variable, following prior work [@long2021parallel]; however, we found that these automated recognition scores were only modestly correlated with human recognition scores ($r$ = `r round(human_v_model$estimate,2)`, $t$ = `r round(human_v_model$estimate,2)`, p < .001); furthermore, descriptive plots of the automated recognition scores revealed a difference between the two drawing tasks that was not evident in the human data. To be conservative, we thus fit all of our models to the human recognition data but have made all model classifications available for future researcs.

@@

# Discussion
Here, we examined the consistency and variability in the developmental trajectory of children's drawings across two different drawing tasks and geographical locations. First, we found that asking children to draw from memory ("Can you draw a [cat]?") vs. from perception ("Can you draw this [cat] as it looks in the picture?") had little effect on the recognizability of the drawings that children produced. While older children may have integrated specific visual information that was present in the photographs, in keeping with prior work [@bruner1984], this did not seem to change how recognizable their drawings were to outside observers. Overall, these results suggest that younger children's drawings are unlikely to be constrained by their ability to recall specific, canonical exemplars of object categories when they are trying to draw them.

>>>>

### Automated recognition scores
We used a combination of deep CNN activations and logistic regressions to obtain automated recognition scores, as per our pre-registered protocol. To encode the high-level visual features of each sketch, we used the VGG-19 architecture [@simonyan2014very] a deep convolutional neural network pre-trained on Imagenet classification. We used model activations in the second-to-last layer of this network, which is the first fully connected layer of the network (FC6). Raw feature representations in this layer consist of flat 4096-dimensional vectors, to which we applied channel-wise normalization across all filtered drawings in the dataset.  Next, we used these features to train object category decoders. We then trained a 12-way logistic classifier with L2 regularization (tolerance = .1, regularization = .1), and used this classifier to estimate the category label for each drawing in the dataset.To avoid any bias due to imbalance in the distribution of drawings over categories, we randomly under sampled such that there were an equal number of drawings for each combination of geographical site (San Jose, Beijing) and the 12 categories. No additional metadata about the age of the child who produced each sketch was provided to the decoder. This procedure was repeated for each drawing in the dataset, yielding a binary a recognition score for each drawing.

@@

This analysis plan was pre-registered at https://osf.io/qymjr/; however, we had initially planned to use automated recognition scores as our main dependent variable, following prior work [@long2021parallel]; however, we found that these automated recognition scores were only modestly correlated with human recognition scores ($r$ = `r round(human_v_model$estimate,2)`, $t$ = `r round(human_v_model$statistic,2)`, p < .001); furthermore, descriptive plots of the automated recognition scores revealed a difference between the two drawing tasks that was not evident in the human data. To be conservative, we thus fit all of our models to the human recognition data but have made all model classifications available for future researcs.

@@

# Discussion
Here, we examined the consistency and variability in the developmental trajectory of children's drawings across two different drawing tasks and geographical locations. First, we found that asking children to draw from memory ("Can you draw a [cat]?") vs. from perception ("Can you draw this [cat] as it looks in the picture?") had little effect on the recognizability of the drawings that children produced. While older children may have integrated specific visual information that was present in the photographs, in keeping with prior work [@bremmer1984prior], this did not seem to change how recognizable their drawings were to outside observers. Overall, these results suggest that younger children's drawings are unlikely to be constrained by their ability to recall specific, canonical exemplars of object categories when they are trying to draw them.
